---
title: "Reproductive Number of TB in the US - Finals Results Plots (For Main Paper)"
author: "Kristina Yamkovoy"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

Note: these results are for the main paper. This file does not contain supplement plots.

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(readxl)
library(openxlsx)

output_folder = '../output_from_cdc/output_23nov20/output/'
```

# Country-level estimates

## Overall US

### Line plot with and without importation
```{r}
us = read.csv('../output_from_cdc/output_10jul20/output/overall_us.csv', stringsAsFactors = FALSE)
us$date = as.Date(paste(us$date, '-01', sep=''))

us_import = read.csv('../output_from_cdc/output_10jul20/output/overall_us_import.csv', stringsAsFactors = FALSE)
us_import$date = as.Date(paste(us_import$date, '-01', sep=''))

us$import = 'no'
us_import$import = 'yes'

us_2 = rbind(us, us_import)

us_2$import = factor(us_2$import, levels = c('yes','no'))

# note: cut off first year and last two years
ggplot(us_2, aes(date, rts, color = import)) +
  geom_line(size = 1.1) +
  scale_y_continuous(limits = c(0.75, 1.5), expand = c(0,0)) +
  scale_x_date(limits = c(as.Date('2011-01-01'), as.Date('2016-01-01')), expand = c(0, 0), 
               date_breaks = '1 year', date_labels = "%Y") +
  #ggtitle('Reproductive Number of TB in United States') +
  xlab('Year') + ylab(expression('R'[t])) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = import), alpha = 0.3) +
  scale_fill_manual(values = c('grey20', 'grey60'), name = 'Incorporating time since \narrival to the U.S.?',
                    labels = c('Yes', 'No')) +
  scale_color_manual(values = c('grey20', 'grey60'), name = 'Incorporating time since \narrival to the U.S.?',
                     labels = c('Yes', 'No'))+
  #scale_fill_grey(name = 'Importation', labels = c('No', 'Yes')) +
  #scale_color_grey(name = 'Importation', labels = c('No', 'Yes')) +
  #scale_fill_brewer(name = 'Importation', labels = c('No', 'Yes'), palette = "Set1") +
  #scale_color_brewer(name = 'Importation', labels = c('No', 'Yes'), palette = "Set1") +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()
```

Note: do not include boxplots for overall US - can mention all time average in text with and without incorporating importation.

## Origin of birth
```{r}
us = read.csv('../output_from_cdc/output_04dec20/output/overall_us_usborn_import_ci_all.csv', stringsAsFactors = FALSE)
us = us[us$usborn != 'UNK', ]

ggplot(us, aes(x = usborn, group = usborn, fill = usborn)) +
  geom_boxplot(aes(ymin = min, lower = q25, middle = rt_est, upper = q75, ymax = max), stat = "identity") + 
  xlab('Origin of Birth') + ylab(expression('R'[t])) +
  scale_x_discrete(labels = c('Foreign Born', 'US Born')) +
  scale_fill_manual(values = c('grey50', 'grey70')) +
  #ggtitle('Origin of Birth') +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
```


## Age group {.tabset}

### No interaction
```{r}
us = read.csv(paste(output_folder, 'overall_us_age_import_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
us = us[us$age != 'UNK', ]

ggplot(us, aes(x = age, y = rt_est)) +
  geom_point(position = position_dodge(width=0.5), size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.4, position = position_dodge(width=0.5)) +
  xlab('Age (years)') + ylab(expression('R'[t])) +
  #ggtitle('Age - No Interaction') +
  scale_x_discrete(labels = c('0-4', '5-14', '15-24', '25-44', '45-64', '65+')) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()

```


### Interaction (POLYMOD)
```{r}
us = read.csv(paste(output_folder, 'overall_us_age_import_polymod_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
us = us[us$age != 'UNK', ]

ggplot(us, aes(x = age, y = rt_est)) +
  geom_point(position = position_dodge(width=0.5), size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.4, position = position_dodge(width=0.5)) +
  xlab('Age (years)') + ylab(expression('R'[t])) +
  #ggtitle('Age - Interaction (POLYMOD)') +
  scale_x_discrete(labels = c('0-4', '5-14', '15-24', '25-44', '45-64', '65+')) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()
```


## US born with age {.tabset}

### No interaction
```{r}
usborn = read.csv(paste(output_folder, 'overall_us_age_import_usborn_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
usborn = usborn[usborn$age != 'UNK', ]
usborn$oob = 'US Born'

forborn = read.csv(paste(output_folder, 'overall_us_age_import_notusborn_ci_all.csv', sep = ''),
                   stringsAsFactors = FALSE)
forborn = forborn[forborn$age != 'UNK', ]
forborn$oob = 'Foreign Born'

us_oob = rbind(usborn, forborn)


ggplot(us_oob, aes(x = age, y = rt_est, group = oob, shape = oob)) +
  geom_point(position = position_dodge(width=0.5), size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.4, position = position_dodge(width=0.5)) +
  xlab('Age (years)') + ylab(expression('R'[t])) +
  #ggtitle('Age and Origin of Birth (Ver 2) - No Interaction') +
  #scale_shape_manual(name = 'Origin of Birth', values = c(8,4)) +
  scale_x_discrete(labels = c('0-4', '5-14', '15-24', '25-44', '45-64', '65+')) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()

```


### Interaction (POLYMOD)
```{r}
usborn = read.csv(paste(output_folder, 'overall_us_age_import_usborn_polymod_ci_all.csv', sep = ''),
                  stringsAsFactors = FALSE)
usborn = usborn[usborn$age != 'UNK', ]
usborn$oob = 'US Born'

forborn = read.csv(paste(output_folder, 'overall_us_age_import_notusborn_matrix_ci_all.csv', sep = ''),
                   stringsAsFactors = FALSE)
forborn = forborn[forborn$age != 'UNK', ]
forborn$oob = 'Foreign Born'

us_oob = rbind(usborn, forborn)


ggplot(us_oob, aes(x = age, y = rt_est, group = oob, shape = oob)) +
  geom_point(position = position_dodge(width=0.5), size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.4, position = position_dodge(width=0.5)) +
  xlab('Age (years)') + ylab(expression('R'[t])) +
  #ggtitle('Age and Origin of Birth (Ver 2) - Interaction (POLYMOD)') +
  scale_shape_manual(name = 'Origin of Birth', values = c(8,4)) +
  scale_x_discrete(labels = c('0-4', '5-14', '15-24', '25-44', '45-64', '65+')) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()
```


# Region-level estimates

## All {.tabset}

### Heatmap
```{r}
library(usmap)

# states in each region
reg_1 = c('CT', 'ME', 'MA', 'NH', 'RI', 'VT')
reg_10 = c('ID', 'OR', 'WA', 'AK')
reg_2 = c('NJ', 'NY')
reg_3 = c('DE', 'MD', 'PA', 'VA', 'WV')
reg_4 = c('AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN')
reg_5 = c('IL', 'IN', 'MI', 'MN', 'OH', 'WI')
reg_6 = c('AR', 'LA', 'NM', 'OK', 'TX')
reg_7 = c('IA', 'KS', 'MO', 'NE')
reg_8 = c('CO', 'MT', 'ND', 'SD', 'UT', 'WY')
reg_9 = c('AZ', 'CA', 'NV', 'HI')

reg_all = read.csv(paste(output_folder, 'region_import_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
reg_all$num_state = c(6, 4, 2, 5, 8, 6, 5, 4, 6, 4)

reg_repeat = reg_all %>%
  group_by(region) %>%
  slice(rep(1:n(), each = num_state))

reg_repeat$state = c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'ID', 'OR', 'WA', 'AK', 'NJ', 'NY',
                     'DE', 'MD', 'PA', 'VA', 'WV', 'AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN',
                     'IL', 'IN', 'MI', 'MN', 'OH', 'WI', 'AR', 'LA', 'NM', 'OK', 'TX',
                     'IA', 'KS', 'MO', 'NE', 'CO', 'MT', 'ND', 'SD', 'UT', 'WY', 'AZ', 'CA', 'NV', 'HI')


plot_usmap(data = reg_repeat, values = "rt_est") +
  scale_fill_gradient(name = expression('R'[t]), high = "#132B43", low = "#56B1F7") +
  theme(legend.position = "right")

```

### Boxplots
```{r}
us = read.csv(paste(output_folder, 'region_import_ci_all.csv', sep = ''), stringsAsFactors = FALSE)

ggplot(us, aes(x = region, group = region, fill = region)) +
  geom_boxplot(aes(ymin = min, lower = q25, middle = rt_est, upper = q75, ymax = max), stat = "identity") + 
  xlab('Region') + ylab(expression('R'[t])) +
  #ggtitle('Region') +
  scale_fill_manual(values = rep('grey80', 10)) +
  scale_x_discrete(name = 'HHS Region', labels = 1:10) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
```


## Origin of birth and region {.tabset}

### Heatmaps

#### US Born
```{r}
reg_us = read.csv(paste(output_folder, 'region_usborn_import_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
reg_us$num_state = c(6, 4, 2, 5, 8, 6, 5, 4, 6, 4)

reg_repeat_us = reg_us %>%
  group_by(region) %>%
  slice(rep(1:n(), each = num_state))

reg_repeat_us$state = c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'ID', 'OR', 'WA', 'AK', 'NJ', 'NY',
                     'DE', 'MD', 'PA', 'VA', 'WV', 'AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN',
                     'IL', 'IN', 'MI', 'MN', 'OH', 'WI', 'AR', 'LA', 'NM', 'OK', 'TX',
                     'IA', 'KS', 'MO', 'NE', 'CO', 'MT', 'ND', 'SD', 'UT', 'WY', 'AZ', 'CA', 'NV', 'HI')


plot_usmap(data = reg_repeat_us, values = "rt_est") +
  scale_fill_gradient(name = expression('R'[t]), high = "#132B43", low = "#56B1F7") +
  theme(legend.position = "right")

```

#### Foreign Born
```{r}
reg_for = read.csv(paste(output_folder, 'region_notusborn_import_ci_all.csv', sep = ''), stringsAsFactors = FALSE)
reg_for$num_state = c(6, 4, 2, 5, 8, 6, 5, 4, 6, 4)

reg_repeat_for = reg_for %>%
  group_by(region) %>%
  slice(rep(1:n(), each = num_state))

reg_repeat_for$state = c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'ID', 'OR', 'WA', 'AK', 'NJ', 'NY',
                     'DE', 'MD', 'PA', 'VA', 'WV', 'AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN',
                     'IL', 'IN', 'MI', 'MN', 'OH', 'WI', 'AR', 'LA', 'NM', 'OK', 'TX',
                     'IA', 'KS', 'MO', 'NE', 'CO', 'MT', 'ND', 'SD', 'UT', 'WY', 'AZ', 'CA', 'NV', 'HI')


plot_usmap(data = reg_repeat_for, values = "rt_est") +
  scale_fill_gradient(name = expression('R'[t]), high = "#132B43", low = "#56B1F7") +
  theme(legend.position = "right")

```


### Boxplots
```{r}
usborn = read.csv(paste(output_folder, 'region_usborn_import_ci_all.csv', sep = ''),
                  stringsAsFactors = FALSE)
usborn$oob = 'US Born'

forborn = read.csv(paste(output_folder, 'region_notusborn_import_ci_all.csv', sep = ''),
                   stringsAsFactors = FALSE)
forborn$oob = 'Non-US Born'

us_oob = rbind(usborn, forborn)

us_oob$oob = factor(us_oob$oob, levels = c('US Born', 'Non-US Born'))
us_oob$region = factor(us_oob$region, levels = c('reg_1', 'reg_2', 'reg_3', 'reg_4', 'reg_5',
                                                 'reg_6', 'reg_7', 'reg_8', 'reg_9', 'reg_10'),
                                      labels = seq(1,10))

ggplot(us_oob, aes(x = region, group = interaction(region, oob), fill = oob)) +
  geom_boxplot(aes(ymin = min, lower = q25, middle = rt_est, upper = q75, ymax = max), stat = "identity") + 
  xlab('HHS Region') + ylab(expression('R'[t])) +
  #ggtitle('Region and Origin of Birth (Ver 1)') +
  scale_fill_manual(name = "Origin of Birth", values = c('grey60', 'grey95')) +
  #scale_x_discrete(labels = c(1, 10, 2:9)) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw() 
  # theme(panel.grid.minor.x = element_blank(),
  #       panel.grid.major.x = element_blank())


ggplot(us_oob, aes(x = region, y = rt_est, group = oob, shape = oob)) +
  geom_point(position = position_dodge(width=0), size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.4, position = position_dodge(width=0)) +
  xlab('HHS Region') + ylab(expression('R'[t])) +
  #ggtitle('Region and Origin of Birth (Ver 2)') +
  scale_shape_manual(name = 'Origin of Birth', values = c(15,2), labels = c('U.S.-born', 'Non-U.S.-born')) +
  #scale_x_discrete(labels = c(1, 10, 2:9)) +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  theme_bw()
```


# HHS Regions
```{r}
reg_repeat$reg_factor = factor(reg_repeat$region, levels = c('reg_1', 'reg_2', 'reg_3', 'reg_4', 'reg_5',
                                                             'reg_6', 'reg_7', 'reg_8', 'reg_9', 'reg_10'),
                               labels = 1:10)


plot_usmap(data = reg_repeat, values = "reg_factor") +
  scale_fill_manual(name = "Region", values = c('grey40', 'grey60', 'grey90', 'grey70', 'grey30',
                                                'grey80', 'grey10', 'grey95', 'grey50', 'grey20')) +
  theme(legend.position = "right")

plot_usmap(data = reg_repeat, values = "reg_factor") +
  # scale_fill_manual(name = "Region", values = c('lightseagreen', 'lightsalmon', 'mediumorchid', 'royalblue',
  #                                               'springgreen',
  #                                               'red', 'grey70', 'grey80', 'grey90', 'grey100')) +
  scale_fill_brewer(name = "Region", palette = "Set3", na.translate = F) +
  theme(legend.position = "right")

```


# HHS Regions - Individual
```{r}
w = 1920/2
h = 1080/2

png('reg_1.png', width = w, height = h)
plot_usmap('states', include = c(reg_1, 'CA', 'TX'))
dev.off()

png('reg_2.png', width = w, height = h)
plot_usmap('states', include = c(reg_2, 'CA', 'ME', 'TX'))
dev.off()

png('reg_3.png', width = w, height = h)
plot_usmap('states', include = c(reg_3, 'CA', 'ME', 'TX'))
dev.off()

png('reg_4.png', width = w, height = h)
plot_usmap('states', include = c(reg_4, 'CA', 'ME', 'TX'))
dev.off()

png('reg_5.png', width = w, height = h)
plot_usmap('states', include = c(reg_5, 'CA', 'ME', 'TX'))
dev.off()

png('reg_6.png', width = w, height = h)
plot_usmap('states', include = c(reg_6, 'CA', 'ME', 'TX'))
dev.off()

png('reg_7.png', width = w, height = h)
plot_usmap('states', include = c(reg_7, 'CA', 'ME', 'TX'))
dev.off()

png('reg_8.png', width = w, height = h)
plot_usmap('states', include = c(reg_8, 'CA', 'ME', 'TX'))
dev.off()

png('reg_9.png', width = w, height = h)
plot_usmap('states', include = c(reg_9[1:3], 'ME', 'TX'))
dev.off()

png('reg_10.png', width = w, height = h)
plot_usmap('states', include = c(reg_10[1:3], 'CA', 'ME', 'TX'))
dev.off()

png('reg_9_HI.png', width = w, height = h)
plot_usmap('states', include = c(reg_9[4], 'CA', 'ME', 'TX'))
dev.off()

png('reg_10_AK.png', width = w, height = h)
plot_usmap('states', include = c(reg_10[4], 'CA', 'ME', 'TX'))
dev.off()
```